---
title: "Stat1378 Project Part 1"
author: "Angus Freeman"
date: "03/11/2021"
output: html_document
params:
  data: "project.csv"
bibliography: A3referencelist.bib
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```
```{r echo = FALSE, warning = FALSE, message = FALSE, output = FALSE}
library(tidyverse)
```
```{r echo = FALSE, warning = FALSE, message = FALSE}
mydata <- read.csv(params$data)
```
## Using statistical tests to answer research questions about a group of 1000 individuals.
&nbsp;
 
## 1 Abstract 

We obtained a data set with observations of 1000 individuals aged 26-45, with information on their gender, height, weight and amount of physical activity. We investigated whether certain relationships exist between some of these variables, and found a significant correlation between height and weight using a linear model, that men were statistically taller than women using a t-test, and that there was no association between gender and amount of physical activity.

&nbsp;

## 2 Introduction

We were interested in testing the following relationships:

  1. Is there a linear relationship between height and weight?
  2. Is the mean height of males and females the same?
  3. Is there any association between gender and the amount of physical activity?
  
We investigated these questions in a cohort of 1000 individuals ages 26-45 with whom we had information on their gender, height, weight and physical activity.

&nbsp;

## 3 Methods

We obtained a data set with observations from 1000 individuals aged 26-45 with information on their gender (male, female), height (in cm), weight (in kg), and amount of physical activity (none, moderate, intense).

### Linear Model

To determine if there exists a relationship between height and weight, we performed a linear regression using the following model:

&nbsp;
<center>
Height~$i$~ = $\alpha$ + $\beta$ * Weight~$i$~ + $\epsilon$~$i$~
</center>
&nbsp;

where $\epsilon$~$i$~ ~ $N$(0,$\sigma^2$), and $i$ is from 1, ..., 1000. To test for significance of the relationship we used the following test statistic:

&nbsp;
<center>
$t$ = $\frac{\hat{\beta}}{SE}$
</center>
&nbsp;

where $\hat{\beta}$ is the estimated slope coefficient and SE is the standard error.

&nbsp;

@LMBook

### T-test

To determine if the mean height of males and females was the same, we used a two sample, equal variance t-test with the following test statistic:

&nbsp;
<center>
$t$ = $\frac{\bar{x_{1}} - \bar{x_{2}}}{s_{p}\sqrt{\frac{1}{n_{1}}-\frac{1}{n_2}}}$
</center>
&nbsp;
 
where $\bar{x}$ is the sample mean of the population, $s_{p}$ is the pooled variance, and n is the number of observations in the sample.

&nbsp;

@TTestArticle

### Chi-Squared Test

To determine if there was any association between gender and physical activity, we used a chi-squared test with no continuity correction, with the following chi-square statistic:

&nbsp;
<center>
$\sum_{i = 1}^{n}$$\frac{(O-E)^2}{E}$
</center>
&nbsp;

where n is the number of observations, O is the observed value, and E is the expected value.


In order to perform the necessary tests for these research questions, the Project1378 package was used, available from: https://github.com/AngusFreeman/Project1378.

```{r echo = FALSE, warning = FALSE, message = FALSE, output = FALSE}
#devtools::install_github("AngusFreeman/Project1378") to install from github
library(Project1378)
```

@ChiBook

## 4 Results

### Linear Model
```{r echo = FALSE, warning = FALSE, message = FALSE}
lmassumptions(mydata)
```
1. The first graph shows that it is likely that the relationship can be described by a linear equation.
2. The second graph shows that it is likely that the residuals are independent from each other and have equal variance.
3. The third graph shows that the residuals are normally distributed.

Since all assumptions are met we can assume the linear regression is accurate.
```{r echo = FALSE, warning = FALSE}
lmhypothesis(mydata)
lmfit(mydata)
lmdecision(mydata)
```
### T-Test
```{r echo = FALSE, warning = FALSE}
tassumptions(mydata)
```
1. The graphs show that both samples are approximately normally distributed, and that they have approximately equal variance.
2. The data was collected using a random sampling method, and it was ensured that all observations are independent.

Since all assumptions are met we can assume the t-test is accurate.
```{r echo = FALSE, warning = FALSE}
thypothesis(mydata)
tfit(mydata)
tdecision(mydata)
```
### Chi Squared Test
```{r echo = FALSE, warning = FALSE}
chiassumptions(mydata)
```
1. Gender and Physical activity measured by none, moderate and intense are both categorical variables.
2. The categories of variables are mutually exclusive.
3. The data collection ensured that all observations were independent.
4. The expected value table show that the expected value is greater than 5 for at least 80% of the cells.

Since all assumptions are met we can assume the chi-squared test is accurate.
```{r echo = FALSE, warning = FALSE}
chihypothesis(mydata)
chifit(mydata)
chidecision(mydata)
```



## 5 Conclusion

### Linear Model
```{r echo = FALSE, warning = FALSE}
lmfit(mydata)[4]
```

```{r echo = FALSE, warning = FALSE}
lmconclusion(mydata)
```

### T-Test
```{r echo = FALSE, warning = FALSE}
tfit(mydata)[3]
```

```{r echo = FALSE, warning = FALSE}
tconclusion(mydata)
```

### Chi Squared - Test
```{r echo = FALSE, warning = FALSE}
chifit(mydata)[3]
```

```{r echo = FALSE, warning = FALSE}
chiconclusion(mydata)
```

## 6 References
